Estimating demographic bias on tests of children's early vocabulary

### 2022-02-01 paper/cogsci.tex
minor edits, fig tweaks

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{CSLReferences}%
  {}%
  {\par}

@@

One challenge in comprehensive assessment of these differences is that
is difficult to get a complete measure of young children's language
skill. Long recordings are prohibitively difficult to collect and
transcribe, and yet any short recording (e.g., a 1-hour play session)
will elicit only a small proportion of the words and constructions that
children know. Thus, researchers of early word learning have constructed
tests with hundreds of words, intentionally oversampling words that are
more likely to be known by young children.

We focus here on the MacArthur-Bates Communicative Development
Inventories {[}CDIs; Larry Fenson et al. (1994); L. Fenson et al.
(2007){]}, a set of parent-reported measures of children's productive
and receptive language skills, which offer a low-cost and reliable way
to estimate children's early language skills (Larry Fenson et al.,
1994). CDIs have shown good predictive validity (Bornstein \& Putnick,
2012; Duff, Reen, Plunkett, \& Nation, 2015; e.g., Larry Fenson et al.,
1994). Our primary focus is the vocabulary checklist portion of the CDI
Words \& Sentences (CDI:WS) form, comprised of 680 early-learned words
across 22 categories (e.g., animals, vehicles, action words, pronouns)
selected to assess the productive vocabulary of children 16 to 30 months
of age. For each item on the CDI:WS, caregivers are asked to respond
whether the target child has been heard to say (i.e.~produce) the given
item. Children's total vocabulary score on the CDI:WS is tightly
correlated with other facets of early language (e.g., grammatical
competence and gesture), suggesting that the language system is
``tightly woven'' (Frank, Braginsky, Yurovsky, \& Marchman, 2021). Due
to these desirable properties, CDIs have been adapted to dozens of
languages, and a central repository of CDI data contributed from all
over the world has been created {[}Wordbank; Frank, Braginsky, Yurovsky,
\& Marchman (2017); Frank, Braginsky, Yurovsky, \& Marchman (2021){]}.

@@

The idea that some items on a test may show bias, favoring one group
over another, is known in psychometrics as Differential Item Function
(DIF; Holland \& Wainer (1993){]}. DIF can decrease the validity of a
test, as the test may be overestimating the ability of test-takers in
one group and underestimating the ability of another group -- not
because of any underlying mean difference in ability between the groups,
but simply because the test is unfair (Camilli, 2013). For example, if a
vocabulary test is composed almost entirely of the names of farm
equipment, the test will underestimate the knowledge of urban children,
who have experience of other contexts, despite being less familiar with
farm equipment than their rural peers. Of course, there may be an
ability difference between demographic groups, as is likely the case for
the female language advantage, which has been documented in many
languages (Eriksson et al., 2012; Frank, Braginsky, Yurovsky, \&
Marchman, 2021). And both observations could be true -- a test could be
biased to inflate a true demographic difference.

@@

We analyzed parent-reported Wordbank data from 5520 American English
CDI: Words \& Sentences administrations for children 16 to 30 months of
age (Frank, Braginsky, Yurovsky, \& Marchman, 2017, 2021). Full
demographic data are not reported in some datasets contributed to
Wordbank: sex was available for 4094 children, race/ethnicity was
available for 2715 children, and maternal education (a proxy for
socioeconomic status; SES) was available for 5520 children.

The analysis of sex-based differences included CDI administrations from
1989 female and 2105 male children. The analysis of race-based
differences included data from 2202 white, 67 Asian, 222 Black, 131
Hispanic, and 93 ``Other'' children. Due to sparse data for many
categories, we binarized participants' race andethnicity as White (2202)
or Non-white (513), recognizing that there is important variation
between groups that this will fail to capture. Data for the maternal
education analysis included CDI data from children whose mothers had the
following levels of education: 8 with no more than primary school
education, 123 with some secondary school, 416 with no more than
secondary school, 613 with some college, 870 with no more than a college
degree, 162 with no more than some graduate school, and 584 with a
graduate degree. Again due to data sparsity, we binarized the 4973
children whose mothers had at least some college or more as high
maternal education (high-ME), and those whose mothers had at most high
school (547 children) as low maternal education (low-ME).

@@

First, we examine the size of demographic effects on language ability in
a baseline Rasch model fitted without regard to demographic group. Then
we fit a multigroup Rasch model for each demographic factor, and
characterize the between-group differences in item difficulties. Next,
we re-examine the size of demographic effects after pruning biased items
from the CDI:WS using different criteria, and recommend a threshold.
Finally, we measure the strength of association between sex-related
differences in language input and the degree of sex-related DIF.

@@

Figure 2 shows GLIMMERs for a selection of CDI items for sex (left),
maternal education (middle), and race (right). The full GLIMMERs, with
all 680 CDI:WS items, are available on
\href{https://osf.io/57rsw/?view_only=2b6ecb61fe08458293af7421d276932a}{OSF},
but were too large to include here. It is important, however, to inspect
the full plots, for if there is a cluster of items in a GLIMMER, the
analyst may conclude that these items are strong candidates for DIF on
that dimension. For example, there is some clustering at the top and
bottom of Figure 2: at the top, ``vagina''\footnote{Parents using the
  CDI are instructed to check genital items if the child produces
  whatever word the family uses.}, ``tights,'' ``dress (object),'' and
``doll'' form a cluster of items that are much more well-known for
females, while at the bottom, ``penis'' stands out as much more
well-known by males. For the maternal education and race GLIMMERs
(Figures 3 and 4) -- and in the rest of the full sex GLIMMER, there are
not clusters, but rather a continuum of smoothly varying differences
with overlap. This pattern makes identifying items with DIF quite
difficult, as different methods are likely to yield inconsistent results
(Stenhaug, Frank, \& Domingue, 2021). Hence, we next characterized the
distributions of the item-level group difficulty differences, and
measured the influence of pruning a varying number of items on the
demographic effect sizes.

@@

\caption[GLIMMER plot of a sample of CDI:WS words from the sex bias model]{GLIMMER plot of a sample of CDI:WS words from the sex bias model. Words at the top are more well-known by females, while those at the bottom are more known for males.}\label{fig:glimmer-combo}
\end{figure*}
\end{CodeChunk}

@@

For sex, the median difficulty difference (male-female) was 0.28
(M=0.27, sd=0.4), with 593/680 items being easier for females than
males. The fact that the bulk of this distribution favors females shows
that the female language advantage is pervasive across the CDI:WS, and
suggests that it is likely to reflect more than the effect of a small
number of biased words (cf. Frank, Braginsky, Yurovsky, \& Marchman,
2021).

@@

Figure 4 shows the 59 items showing extreme bias (i.e., whose difficulty
was \textgreater2.25 SD from the mean difficulty difference on one or
more demographic dimension). There were 22 items with extreme sex-based
difficulty differences, only 7 of which were more known by females, with
the other 15 items favoring males. All but one of the sex-biased items
are nouns, many of which are stereotypically associated with one gender
more than the other, including stereotypically-male professions (e.g.,
``fireman''). For maternal education, 27 extrema were identified, only
10 of which were biased for low-SES children. As for sex, most of the
ME-biased items are nouns (15/27). Notably, many of the items more known
by high-ME children were animals (``zebra,'' ``owl'') and animal sounds
(``baa baa,'' ``moo''). Items more known by low-ME children include a
few sweet treats: ``candy,'' ``soda/pop,'' and ``gum.'' For race, 21
extrema were identified, only 8 of which were easier for non-white
children. The lexical class of the race-biased items was more varied: 10
were nouns, but many others were early-learned words and phrases (e.g.,
``up,'' ``all gone,'' ``uh-oh,'' ``bye,'' ``mommy''). Only 10 of the 59
items were extreme in more than one model, with most of the overlap
being between race and ME (8 items). Only 1 item (``vroom'') was extreme
for all three demographics.

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics[width=\linewidth]{figs/extrema2p25sd}

@@

\caption[The 59 items showing extreme bias (difficulty difference $>2.25$ SD) for one or more demographic]{The 59 items showing extreme bias (difficulty difference $>2.25$ SD) for one or more demographic.}\label{fig:biased-words}
\end{figure}
\end{CodeChunk}

@@

We investigated the CDI:WS, a popular parent-report measure of
children's early vocabulary, for potential demographic bias, examining
the distribution of word difficulties for children of mothers with high
vs.~low education (a proxy for family SES), for female vs.~male
children, and for white vs.~non-white children. Our IRT-based analysis
revealed differential item functioning (DIF) for many items along each
demographic dimension, but only in the case of sex were clear clusters
of items that were more well-known to females (including feminine
clothing and genitalia), and a clear item that was more well-known to
males (male genitalia). For the rest of the items, and for SES- and
race-based analysis, there was a smooth continuum of DIF, making the
boundary of true DIF subjective. To move forward, we identified
candidate DIF items by systematically pruning the extremes of each
distribution, excluding outliers for each demographic across a range of
thresholds, and seeking to minimize the size of demographic effects.
Although many exclusion thresholds decreased the size of maternal
education and race effects, the female advantage actually grew under
most prunings, with the majority of excluded extrema being
stereotypically-male nouns (e.g, ``truck,'' ``fireman'').

This analysis highlights a fundamental difficulty of identifying DIF: an
analyst must either know the expected magnitude of an ability difference
between the two demographic groups, or know a set of ``anchor'' items
that are equally difficult (i.e., unbiased) for both groups (for an
overview, see Stenhaug, Frank, \& Domingue, 2021). But in the universe
of children's early language, there is only a finite set of
early-learned words to choose from -- and we may expect many of them to
be biased for various environmental reasons (e.g., boys and girls know
the names for their respective anatomy; children in Florida may not use
mittens or skis). Hence, the presence of DIF on a set of items could
indicate varying linguistic and environmental input, rather than an
actual advantage for one demographic over another.

@@

This investigation is only a first step in measuring demographic bias in
the items on the CDI:WS. More research is needed to determine whether
CDI items that show DIF are more related to environmental frequencies,
as we found only a modest association between the relative frequency of
words in child-directed speech to boys vs.~girls and the amount of
sex-based DIF. Examining the relative frequency of CDI words across
households with varying maternal education and race may reveal similarly
weak associations, or there may be strong associations only among
particular categories of words. It may be that particular activity
contexts--popular with some demographic groups, and not others--may be
predictive. For example, many of the extrema favoring high-ME children
are animals and animal sounds (e.g., ``quack quack,'' ``woof woof,''
``baa baa,'' ``duck,'' ``sheep,'' ``giraffe,'' ``zebra''): do high-SES
households visit the zoo more often, or perhaps engage in other
activities related to naming animals and noises they make (e.g., animal
noise toys)? If certain activities are driving early word learning for
high-ME children, what are the activities (and associated vocabulary)
that low-SES households are instead engaging in? Critically this
enterprise will require a greater diversity of publicly-available
corpora of child language; we view the lack of such data as a
substantial issue.

SES, race, and sex are only three of many possible demographic
dimensions of interest. For example, geographic region is likely
predictive of children's early experience with (and thus knowledge of)
CDI items related to winter weather (e.g., ``snow,'' ``mittens'') or
outdoors activities (e.g.~``camping''). A truly fair test of children's
early vocabulary would contain a representative and balanced sample of
words from all activities that children engage in, across demographic
groups.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-bornstein2016stability}{}%
Bornstein, M. H., Hahn, C.-S., \& Putnick, D. L. (2016). Stability of
core language skill across the first decade of life in children at
biological and social risk. \emph{Journal of Child Psychology and
Psychiatry}, \emph{57}(12), 1434--1443.

@@

\leavevmode\hypertarget{ref-R-mirt}{}%
Chalmers, R. P. (2012). {mirt}: A multidimensional item response theory
package for the {R} environment. \emph{Journal of Statistical Software},
\emph{48}(6), 1--29.
http://doi.org/\href{https://doi.org/10.18637/jss.v048.i06}{10.18637/jss.v048.i06}

@@

\leavevmode\hypertarget{ref-fenson1994}{}%
Fenson, Larry, Dale, P. S., Reznick, J. S., Bates, E., Thal, D. J.,
Pethick, S. J., \ldots{} Stiles, J. (1994). Variability in early
communicative development. \emph{Monographs of the Society for Research
in Child Development}, i--185.

\leavevmode\hypertarget{ref-Fenson2007}{}%
Fenson, L., Marchman, V. A., Thal, D. J., Dale, P. S., Reznick, J. S.,
\& Bates, E. (2007). \emph{{M}ac{A}rthur-{B}ates {C}ommunicative
{D}evelopment {I}nventories: User's guide and technical manual (2nd
ed.)}. Baltimore, MD: Brookes.

@@

\leavevmode\hypertarget{ref-macwhinney2000childes}{}%
MacWhinney, B. (2000). \emph{The CHILDES project: Tools for analyzing
talk. Transcription format and programs} (Vol. 1). Psychology Press.

@@

\leavevmode\hypertarget{ref-rowe2009differences}{}%
Rowe, M. L., \& Goldin-Meadow, S. (2009). Differences in early gesture
explain SES disparities in child vocabulary size at school entry.
\emph{Science}, \emph{323}(5916), 951--953.

\leavevmode\hypertarget{ref-schwab2016}{}%
Schwab, J. F., \& Lew-Williams, C. (2016). Language learning,
socioeconomic status, and child-directed speech. \emph{WIREs Cognitive
Science}, \emph{7}, 264--275.
http://doi.org/\href{https://doi.org/10.1002/wcs.1393}{10.1002/wcs.1393}

>>>>

@@

One challenge in comprehensive assessment of these differences is that
it is difficult to get a complete measure of young children's language
skill. Long recordings are prohibitively difficult to collect and
transcribe, and yet any short recording (e.g., a 1-hour play session)
will elicit only a small proportion of the words and constructions that
children know. Thus, researchers of early word learning have constructed
tests with hundreds of words, intentionally oversampling words that are
more likely to be known by young children.

We focus here on the MacArthur-Bates Communicative Development
Inventories (CDIs; Fenson et al., 1994, 2007), a set of parent-reported
measures of children's productive and receptive language skills, which
offer a low-cost and reliable way to estimate children's early language
skills (Fenson et al., 1994). CDIs have shown good predictive validity
(e.g., Fenson et al., 1994; Bornstein \& Putnick, 2012; Duff, Reen,
Plunkett, \& Nation, 2015). Our primary focus is the vocabulary
checklist portion of the CDI Words \& Sentences (CDI:WS) form, comprised
of 680 early-learned words across 22 categories (e.g., animals,
vehicles, action words, pronouns) selected to assess the productive
vocabulary of children 16 to 30 months of age. For each item on the
CDI:WS, caregivers are asked to respond whether the target child has
been heard to say (i.e.~produce) the given item. Children's total
vocabulary score on the CDI:WS is tightly correlated with other facets
of early language (e.g., grammatical competence and gesture), suggesting
that the language system is ``tightly woven'' (Frank et al., 2021). Due
to these desirable properties, CDIs have been adapted to dozens of
languages, and a central repository of CDI data contributed from all
over the world has been created (Wordbank; Frank, Braginsky, Yurovsky,
\& Marchman, 2017; Frank et al., 2021).

@@

The idea that some items on a test may show bias, favoring one group
over another, is known in psychometrics as Differential Item Function
(DIF; Holland \& Wainer (1993){]}. DIF can decrease the validity of a
test, as the test may be overestimating the ability of test-takers in
one group and underestimating the ability of another group -- not
because of any underlying mean difference in ability between the groups,
but simply because the test is unfair (Camilli, 2013). For example, if a
vocabulary test is composed almost entirely of the names of farm
equipment, the test will underestimate the knowledge of urban children,
who have experience of other contexts, despite being less familiar with
farm equipment than their rural peers. Of course, there may be an
ability difference between demographic groups, as is likely the case for
the female language advantage, which has been documented in many
languages (Eriksson et al., 2012; Frank et al., 2021). And both
observations could be true -- a test could be biased to inflate a true
demographic difference.

@@

We analyzed parent-reported Wordbank data from 5520 American English
CDI: Words \& Sentences administrations for children 16 to 30 months of
age (Frank et al., 2017, 2021). Full demographic data are not reported
in some datasets contributed to Wordbank: sex was available for 4094
children, race/ethnicity was available for 2715 children, and maternal
education (a proxy for socioeconomic status; SES) was available for 5520
children.

The analysis of sex-based differences included CDI administrations from
1989 female and 2105 male children. The analysis of race-based
differences included data from 2202 white, 67 Asian, 222 Black, 131
Hispanic, and 93 ``Other'' children. Due to sparse data for many
categories, we binarized participants' race and ethnicity as White
(2202) or Non-white (513), recognizing that there is important variation
between groups that this will fail to capture. Data for the maternal
education analysis included CDI data from children whose mothers had the
following levels of education: 8 with no more than primary school
education, 123 with some secondary school, 416 with no more than
secondary school, 613 with some college, 870 with no more than a college
degree, 162 with no more than some graduate school, and 584 with a
graduate degree. Again due to data sparsity, we binarized the 4973
children whose mothers had at least some college or more as high
maternal education (high-ME), and those whose mothers had at most high
school (547 children) as low maternal education (low-ME).

@@

First, we examine the size of demographic effects on language ability in
a baseline Rasch model fitted without regard to demographic group. Then
we fit a multigroup Rasch model for each demographic factor, and
characterize the between-group differences in item difficulties. Next,
we re-examine the size of demographic effects after pruning biased items
from the CDI:WS using different criteria, and recommend a threshold.
Finally, we measure the strength of association between sex-related
differences in language input and the degree of sex-related DIF. The
scripts and models needed to reproduce our analyses are available on
\href{https://osf.io/57rsw/?view_only=2b6ecb61fe08458293af7421d276932a}{OSF}.

@@

Figure 2 shows GLIMMERs for a selection of CDI items for sex (left),
maternal education (middle), and race (right). The full GLIMMERs, with
all 680 CDI:WS items, are available on
\href{https://osf.io/57rsw/?view_only=2b6ecb61fe08458293af7421d276932a}{OSF},
but were too large to include here. It is important, however, to inspect
the full plots, for if there is a cluster of items in a GLIMMER, the
analyst may conclude that these items are strong candidates for DIF on
that dimension. For example, there is some clustering at the top and
bottom of the sex GLIMMER (Fig. 2, left): at the top,
``vagina''\footnote{Parents using the CDI are instructed to check
  genital items if the child produces whatever word the family uses.},
``tights'', ``dress (object)'', and ``doll'' form a cluster of items
that are much more well-known for females, while at the bottom,
``penis'' stands out as much more well-known by males. For the maternal
education and race GLIMMERs -- and in the rest of the full sex GLIMMER,
there are not clusters, but rather a continuum of smoothly varying
differences with overlap. This pattern makes identifying items with DIF
quite difficult, as different methods are likely to yield inconsistent
results (Stenhaug et al., 2021). Hence, we next characterized the
distributions of the item-level group difficulty differences, and
measured the influence of pruning a varying number of items on the
demographic effect sizes.

@@

\caption[GLIMMER plots of a sample of CDI:WS words from the sex bias model (left), the maternal education model (center), and the race model (right)]{GLIMMER plots of a sample of CDI:WS words from the sex bias model (left), the maternal education model (center), and the race model (right). Words at the top are more well-known by one group (e.g., females, for sex), while those at the bottom are more known by the other group (e.g., males), and the selection in the middle show roughly the median difficulty difference for those groups.}\label{fig:glimmer-combo}
\end{figure*}
\end{CodeChunk}

@@

For sex, the median difficulty difference (male-female) was 0.28
(M=0.27, sd=0.4), with 593/680 items being easier for females than
males. The fact that the bulk of this distribution favors females shows
that the female language advantage is pervasive across the CDI:WS, and
suggests that it is likely to reflect more than the effect of a small
number of biased words (cf.~Frank et al., 2021).

@@

Figure 4 shows the 59 items showing extreme bias (i.e., whose difficulty
was \textgreater2.25 SD from the mean difficulty difference on one or
more demographic dimension). There were 22 items with extreme sex-based
difficulty differences, only 7 of which were more known by females, with
the other 15 items favoring males. All but one of the sex-biased items
are nouns, many of which are stereotypically associated with one gender
more than the other, including stereotypically-male professions (e.g.,
``fireman''). For maternal education, 27 extrema were identified, only
10 of which were biased for low-SES children. As for sex, most of the
ME-biased items are nouns (15/27). Notably, many of the items more known
by high-ME children were animals (``zebra'', ``owl'') and animal sounds
(``baa baa'', ``moo''). Items more known by low-ME children include a
few sweet treats: ``candy'', ``soda/pop'', and ``gum''. For race, 21
extrema were identified, only 8 of which were easier for non-white
children. The lexical class of the race-biased items was more varied: 10
were nouns, but many others were early-learned words and phrases (e.g.,
``up'', ``all gone'', ``uh-oh'', ``bye'', ``mommy''). Only 10 of the 59
items were extreme in more than one model, with most of the overlap
being between race and ME (8 items). Only 1 item (``vroom'') was extreme
for all three demographics.

\begin{CodeChunk}
\begin{figure}[h]

{\centering \includegraphics[width=\linewidth]{figs/extrema2p25sd_3sd_bold}

@@

\caption[The 59 items showing extreme bias (difficulty difference $>2.25$ SD) for one or more demographic]{The 59 items showing extreme bias (difficulty difference $>2.25$ SD) for one or more demographic. The 19 items in bold are those showing a difference $>3$ SD.}\label{fig:biased-words}
\end{figure}
\end{CodeChunk}

@@

We investigated the CDI:WS, a popular parent-report measure of
children's early vocabulary, for potential demographic bias, examining
the distribution of word difficulties for children of mothers with high
vs.~low education (a proxy for family SES), for female vs.~male
children, and for white vs.~non-white children. Our IRT-based analysis
revealed differential item functioning (DIF) for many items along each
demographic dimension, but only in the case of sex were clear clusters
of items that were more well-known to females (including feminine
clothing and genitalia), and a clear item that was more well-known to
males (male genitalia). For the rest of the items, and for SES- and
race-based analysis, there was a smooth continuum of DIF, making the
boundary of true DIF subjective. To move forward, we identified
candidate DIF items by systematically pruning the extremes of each
distribution, excluding outliers for each demographic across a range of
thresholds, and seeking to minimize the size of demographic effects.
Although many exclusion thresholds decreased the size of maternal
education and race effects, the female advantage actually grew under
most prunings, with the majority of excluded extrema being
stereotypically-male nouns (e.g, ``truck'', ``fireman'').

This analysis highlights a fundamental difficulty of identifying DIF: an
analyst must either know the expected magnitude of an ability difference
between the two demographic groups, or know a set of ``anchor'' items
that are equally difficult (i.e., unbiased) for both groups (for an
overview, see Stenhaug et al., 2021). But in the universe of children's
early language, there is only a finite set of early-learned words to
choose from -- and we may expect many of them to be biased for various
environmental reasons (e.g., boys and girls know the names for their
respective anatomy; children in Florida may not use mittens or skis).
Hence, the presence of DIF on a set of items could indicate varying
linguistic and environmental input, rather than an actual advantage for
one demographic over another.

@@

This investigation is only a first step in measuring demographic bias in
the items on the CDI:WS. More research is needed to determine whether
CDI items that show DIF are more related to environmental frequencies,
as we found only a modest association between the relative frequency of
words in child-directed speech to boys vs.~girls and the amount of
sex-based DIF. Examining the relative frequency of CDI words across
households with varying maternal education and race may reveal similarly
weak associations, or there may be strong associations only among
particular categories of words. It may be that particular activity
contexts--popular with some demographic groups, and not others--may be
predictive. For example, many of the extrema favoring high-ME children
are animals and animal sounds (e.g., ``quack quack'', ``woof woof'',
``baa baa'', ``duck'', ``sheep'', ``giraffe'', ``zebra''): do high-SES
households visit the zoo more often, or perhaps engage in other
activities related to naming animals and noises they make (e.g., animal
noise toys)? If certain activities are driving early word learning for
high-ME children, what are the activities (and associated vocabulary)
that low-SES households are instead engaging in? Critically this
enterprise will require a greater diversity of publicly-available
corpora of child language; we view the lack of such data as a
substantial issue.

SES, race, and sex are only three of many possible demographic
dimensions of interest. For example, geographic region is likely
predictive of children's early experience with (and thus knowledge of)
CDI items related to winter weather (e.g., ``snow'', ``mittens'') or
outdoors activities (e.g.~``camping''). A truly fair test of children's
early vocabulary would contain a representative and balanced sample of
words from all activities that children engage in, across demographic
groups.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-bornstein2016stability}{}%
Bornstein, M. H., Hahn, C.-S., \& Putnick, D. L. (2016). Stability of
core language skill across the first decade of life in children at
biological and social risk. \emph{Journal of Child Psychology and
Psychiatry}, \emph{57}(12), 1434--1443.

@@

\leavevmode\hypertarget{ref-R-mirt}{}%
Chalmers, R. P. (2012). mirt: A multidimensional item response theory
package for the R environment. \emph{Journal of Statistical Software},
\emph{48}(6), 1--29. \url{http://doi.org/10.18637/jss.v048.i06}

@@

\leavevmode\hypertarget{ref-fenson1994}{}%
Fenson, L., Dale, P. S., Reznick, J. S., Bates, E., Thal, D. J.,
Pethick, S. J., \ldots{} Stiles, J. (1994). Variability in early
communicative development. \emph{Monographs of the Society for Research
in Child Development}, i--185.

\leavevmode\hypertarget{ref-Fenson2007}{}%
Fenson, L., Marchman, V. A., Thal, D. J., Dale, P. S., Reznick, J. S.,
\& Bates, E. (2007). \emph{MacArthur-Bates Communicative Development
Inventories: User's guide and technical manual (2nd ed.)}. Baltimore,
MD: Brookes.

@@

\leavevmode\hypertarget{ref-macwhinney2000childes}{}%
MacWhinney, B. (2000). \emph{The childes project: Tools for analyzing
talk. Transcription format and programs} (Vol. 1). Psychology Press.

@@

\leavevmode\hypertarget{ref-rowe2009differences}{}%
Rowe, M. L., \& Goldin-Meadow, S. (2009). Differences in early gesture
explain ses disparities in child vocabulary size at school entry.
\emph{Science}, \emph{323}(5916), 951--953.

\leavevmode\hypertarget{ref-schwab2016}{}%
Schwab, J. F., \& Lew-Williams, C. (2016). Language learning,
socioeconomic status, and child-directed speech. \emph{WIREs Cognitive
Science}, \emph{7}, 264--275. \url{http://doi.org/10.1002/wcs.1393}
